

## Loading and preprocessing the data
First load the data 
```{r}
Data <- read.csv("activity.csv")

```


## What is mean total number of steps taken per day?
1. Use of the aggregate function
```{r}

```


The total number of steps taken per day was calculated by using aggregate function. We report the mean and the median also. 
```{r}
steps.date <- aggregate(steps ~ date, data=Data, FUN=sum)

mean(steps.date$steps)
median(steps.date$steps)

```
Histogram of the total number of steps taken each day
```{r}
hist(steps.date$steps, xlab="Total number  steps per day", main=NULL, col="blue")
```

## What is the average daily activity pattern?
Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis).
We're going to use the aggregate function again.

```{r}
steps.interval <- aggregate(steps ~ interval, data=Data, FUN=mean)
plot(steps.interval, type="l")

```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
steps.interval$interval[which.max(steps.interval$steps)]
```


## Imputing missing values

Number of missing values

```{r}
dim(Data[is.na(Data$steps),])[1]
```

Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

```{r}
Data1 <- merge(Data, steps.interval, by="interval", suffixes=c("",".y"))
missing <- is.na(Data$steps)
Data1$steps[missing] <- Data1$steps.y[missing]
Data1 <- Data1[,c(1:3)]
```
```{r}
steps.date <- aggregate(steps ~ date, data=Data1, FUN=sum)
barplot(steps.date$steps, names.arg=steps.date$date, xlab="date", ylab="steps")
mean(steps.date$steps)
median(steps.date$steps)

```


## Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r}
dtype <- function(date) {
    if (weekdays(as.Date(date)) %in% c("sábado", "domingo")) {
        "weekend"
    } else {
        "weekday"
    }
}
Data1$dtype <- as.factor(sapply(Data1$date, dtype))
```

Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r}
par(mfrow=c(2,1))
for (type in c("weekend", "weekday")) {
    steps.type <- aggregate(steps ~ interval,
                            data=Data1,
                            subset=Data1$dtype==type,
                            FUN=mean)
    plot(steps.type, type="l", main=type)
}
```

